Large Scale Marine Protected Areas

### 2019-01-21 docs/PIPA_Palau.tex
Added manuscript with comments from John

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Displacement of fishing effort by Large Scale Marine Protected Areas},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{chicago}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Displacement of fishing effort by Large Scale Marine Protected
Areas\footnote{Work in progress, do not circulate}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-20}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

@@

\begin{document}
\maketitle
\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. Fishing effort is effectively eliminated
within these protected areas upon implementation. The benefits of
reducing effort have been largely studied, but little empirical works
evaluate what happens to the effort after an MPA is created. The
economic and ecological implications of displacing fishing effort are
not yet fully understood. We use identification of fishing activity via
Automatic Identification Systems (AIS) and causal inference techniques
to provide the first analyses of behavioral changes and spatial
redistribution of tuna purse seiners due to the implementation of a
Large Scale Marine Protected Area in the Pacific Ocean. Our work
provides three main findings: 1) aggregate fishing effort remains
relatively unaffected; 2) Vessels that fished inside the protected area
redistribute to adjacent waters; and 3) We observe a crowding effect for
the first years after implementation. Our results not only provide an
impact evaluation of the effect of LSMPAs on fishing activity, but
provide insights into vessel redistribution dynamics, which may have
ecological and economic implications. As countries continue to implement
LSMPAs as a way to reach the stated 10\% target of ocean protection,
managers should consider how fishing effort will change in space and
through time to ensure that fishing effort is not just displaced
elsewhere, leading to overfishing in adjacent waters. While LSMPAs can
provide a wide range of benefits, their implementation must be
accompained with traditional fisheries management to maximize their
effectiveness.
\end{abstract}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environments by
2020. In an effort to meet this target, the world has seen a rapid
increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by
a small number of Large Scale Marine Protected Areas (LSMPA)
\citet{singleton_2014,boonzaier_2016,alger_2017}). Due to weak property
rights, limited habitat transformation, and potentially lower management
costs, LSMPAs provide an opportunity to safeguard the oceans
\citet{game_2009}. Today, a small number of LSMPAs make up at least 80\%
of the managed areas in the ocean \citet{toonen_2013}. LSMPAs may be
instrumental in helping us achieve the 10\% target by 2020
\citep{mccauley_2016}.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. As with customary MPAs, it is important that we
understand the socioeconomic implications of management interventions.
One issue of particular importance is that of the displacement or
redistribution of fishing effort, which may influence the outcomes of an
MPA \citep{smith_2003}. Theoretical models make different assumptions
about the way in which fishers will reallocate fishing effort after an
area closure. The few empirical works focus on small-scale fisheries and
rarely incorporate spatially explicit data. Despite these efforts,
marine conservation is yet to identify the response in fishing effort
and its management implications.

Recent advances in satellite tracking technologies and near real-time
identification of fishing activity provide us with an opportunity to
tackle increase governance and answer long-lasting questions. Here, we
ask, how do fishers respond to the implementation of LSMPA? After a
closure, and if they continue to fish, where do they reallocate fishing
effort? We use identification of fishing activity via Automatic
Identification Systems (AIS) and causal inference techniques to describe
the behavioral changes and spatial redistribution of the industrial tuna
purse seine fleet due to the implementation of a Large Scale Marine
Protected Areas in the Pacific Ocean.

Our work is novel in the sense that it provides empirical evidence of
the effect of Large Scale Marine Protected Areas in fishing behavior and
distribution and can help guide future interventions as countries march
towards meeting the targeted 10\% of ocean protection. Understanding how
effort is displaced from LSMPAs might provide insights into how the
industrial fleet would react to a high seas closure.

The paper is outlined as follows: Section \ref{background} provides more
information on Large Scale Marine Protected Areas and background on
empirical studies of effort redistribution. An overview of the Nauru
Agreement and associated countries, a description of the fleet that
operates in the region, and a brief history of PIPA is also included.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, section \ref{discussion}
provides an extension of our results to other
LSMPAs\footnote{Not yet. but I think this would be interesting} and
discusses our
results\footnote{I might also include a section, after the background, where I explain the different theoretical models of effort redistribution and behavioral responses}.

\hypertarget{background}{%
\section{Background}\label{background}}

@@

The cutoff at which an MPA is considered to be a LSMPA ranges from areas
larger than 30,000 km\textsuperscript{2} as defined by
\citet{desanto_2013} or areas larger than 250,000 km\textsuperscript{2},
as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows
LSMPAs that meet the later, and that are fully no take. LSMPAs are often
implemented in the pelagic environment, where the dominant human
activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The
early literature on LSMPAs focused on the inherent challenges and
difficulties that come with a pelagic environment. \citet{kaplan_2010}
claimed that very large MPAs would result in excessive opportunity costs
and that these would be difficult to enforce. \citet{game_2009}
suggested that most of the challenges could be overcome with the
incorporation of technology, in what then became known as Dynamic Ocean
Management \citep{maxwell_2015}. The effectiveness and difficulties of
implementing LSMPAs are not the focus of this paper
\footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}.
Instead, we focus on the effect and implications of the ones that
\emph{have been} implemented.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}\label{fig:LSMPAs_map}Large Scale Marine
Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated english name is `Protected Area'.}
\end{figure}

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protects two
thirds of the tagged grey reef sharks by effectively excluding fishing
effort. More recently, \citep{bradley_2018} use similar data to
highlight cases of potential illegal shark fishing \emph{inside} a 2
million km\textsuperscript{2} shark sanctuary. To date, no studies have
evaluated the displacement of fishing effort due to LSMPA
implementation.

Spatial closures of this magnitude are likely to induce changes in
fishers' behavior. Theoretical models of fishing effort redistribution
range from the simplistic assumption that effort inside the bounded
region disappears, to spatially explicit models that reallocate fishing
effort based on habitat characteristics, presence of other vessels, and
expected returns \citep{smith_2003,hilborn_2006}. However, these focus
on the long term optimal equilibrium, and redistribution of fishing
effort may not always be optimal, especially over the first years
\citep{stevenson_2013}.

The empirical research that has been done in customary sized MPAs
suggest that resource users may show idiosyncratic responses. For
example, \citet{stevenson_2013} show that a network of MPAs displaced
fishing effort farther away from ports, resulting in higher
\emph{perceived} costs, and increases in catch per unit effort.
\citet{cabral_2017} analyse the redistribution of fishing and
non-fishing vessels following the implementation of a network of MPAs in
California, and find that commercial dive boats follow a
fishing-the-line pattern, while some fishing boats follow an ideal free
distribution. More recently \citet{elahi_2018} used satellite tracking
data to show that a temporal spatial closure caused trawlers to maintain
effort but apply it more intensively elsewhere, particularly along the
borders and closer to shore. The way in which fishers react to a spatial
closure can have major implications in its outcome
\citep{smith_2003,hilborn_2006} highlighting the need to understand how
fishers react to the implementation of LSMPAs, and fishing effort
changes and is spatially redistributed. However, all these closures took
place within Exclusive Economic Zones, where other regulations exist.
This may not always be the case for LSMPAs, where often the entire EEZ
is converted into a LSMPA, leaving fishers with the options of moving to
the high seas or other countries' EEZs.

\hypertarget{nauru-agreement-and-the-phoenix-island-protected-area}{%
\subsection{Nauru agreement and the Phoenix Island Protected
Area}\label{nauru-agreement-and-the-phoenix-island-protected-area}}

The Nauru Agreement was established in 1982 by a select group of Pacific
island nations to manage their important tuna resources. PNA Members
include Federated States of Micronesia, Kiribati, Marshall Islands,
Nauru, Palau, Papua New Guinea, Solomon Islands, and Tuvalu. The Nauru
Agreement regulated access of foreign vessels (\emph{i.e.} those from
non-PNA countries). Holding \textasciitilde{}80\% of the historical
purse seining grounds within their Exclusive Economic Zones, PNA
countries gained bargaining power when providing access to foreign
fleets \citep{havice_2010}.

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisers recommend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then use or sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. Countries that belong
to the PNA are shown in blue, while empty polygons indicates all others.
A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

@@

One of the most notable and recent management interventions in the
region is the implementation of the Phoenix Island Protected Area (PIPA)
by the government of Kiribati. PIPA was first declared in 2006, and
established in 2008 with only 4\% of it was declared as no-take. In
January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
expanded to a total area of 397,447 km\textsuperscript{2}, roughly 1.5
times the size of Ecuador. Figure \ref{fig:PNA_map} shows a map of the
PNA countries and the Phoenix Island Protected Area.

The closure of such a large area in one of the most important fishing
regions in the world provides a great opportunity to evaluate the
behavioral responses and redistribution of fishing effort by vessels
that used to fish there. PIPA has been the focus of previous research,
showing that fishing effort is effectively reduced after implementation
\citep{mccauley_2016,mcdermott_2018}. To this, we pose two questions:
How do individual vessels respond to the sudden exclusion of such a big
area? And where did all the vessels go? In the next sections we describe
the data and methods used to answer this questions.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

This section is divided into two main parts. First, we provide a general
description of AIS data and the process of identification of
vessel-level fishing events done by Global Fishing Watch\footnote{Global
  Fishing Watch: \url{globalfishingwatch.org}}. Alongside, we describe
the subset of data used on our analyses. When relevant, we also point
out possible shortcomings in the data, or factors that must be
considered in the later analyses. We then move on to explain our
empirical strategy for the identification of the behavioral changes and
redistribution of fishing effort.

\hypertarget{data}{%
\subsection{Data}\label{data}}

Automatic Identification Systems (AIS) are on-board devices that provide
at-sea safety and prevent ship collisions by broadcasting vessel
position, course, and activity to surrounding vessels. These broadcasted
messages can be received by satellites and land-based antennas. GFW uses
convolutional neural networks to infer vessel characteristics and
whether each broadcasted position represents a fishing event, thus
allowing us to estimate near real-time fishing events globally since
2012 \citep{kroodsma_2018}.

The amount of data gathered is dependent on the number of antennas and
satellites that can pick up the signals. In June 1\textsuperscript{st}
2014 satellite count increased from 3 to 6, and then to 10 in January
1\textsuperscript{st} 2016. This causes an increase in the number of
\emph{received} AIS messages (\emph{i.e.} points), and therefore an
apparent increase in the number of vessels and fishing hours. However,
the addition of satellites ``affects'' all vessels in the same way. The
variability in AIS data and ocean conditions require that temporal
trends be taken into account. We do that by including specific controls
in our identification strategy and using a subset of data that meet a
BACI design -which gives us the full tracks for vessels affected and
unaffected by the implementation of PIPA.

Our analysis focuses on purse seine vessels, the most important fishery
for PNA
countries\footnote{We perform some of the same analyses for longliners and include them in the Appendix}.
We identify a total of 103 purse seiners that fished in PNA waters at
least once before
2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses.}.
These vessels represent over 26 million individual observations for the
2012 - 2017 period. We identify 65 vessels that have fished inside PIPA
at least once since 2012. From these, 62 did so before the announcement
(\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}) but are
only able to track 61 for the complete period of study before and after
its
implementation\footnote{1 vessel never fished again after August 18, 2013.}.
On the other hand, 38 vessels never fished inside PIPA but we only have
data for 26 vessels before and after.

>>>>

\documentclass[11pt,english]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Displacement of fishing effort by Large Scale Marine Protected Areas},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{chicago}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{Large Scale Marine Protected Areas}

\author{
  Juan CarlosVillase\~{n}or-Derbez\footnote{Bren School of Environmental Science \& Management, University of California at Santa Barbara, Santa Barbara, CA} \\
  \texttt{juancarlos@ucsb.edu}
  \and
 John Lynham\footnote{Department of Economics, University of Hawaii at Manoa, Honolulu, HI}\\
  \texttt{lynham@hawaii.edu }
}

\date{Updated on: \today}

@@

% John's additional packages below

\usepackage{graphicx}
\graphicspath{{figures/}} %Setting the graphicspath

\begin{document}
\maketitle
\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. Fishing effort is effectively eliminated
within these protected areas upon implementation. The benefits of
reducing effort have been largely studied, but little empirical works
evaluate how vessels react and redistribute after an MPA is created. The
economic and ecological implications of displacing fishing effort are
not yet fully understood. We use identification of fishing activity via
Automatic Identification Systems (AIS) and causal inference techniques
to provide the first analysis of behavioral changes and spatial
redistribution of tuna purse seiners due to the implementation of a
Large Scale Marine Protected Area in the Pacific Ocean. Our work
provides three main findings: 1) aggregate fishing effort remains
relatively unaffected; 2) vessels that fished inside the protected area
redistribute to adjacent waters; and 3) we observe a crowding effect for
the first months after implementation. Our results not only provide an
impact evaluation of the effect of LSMPAs on fishing activity, but
provide insights into vessel redistribution dynamics, which may have
ecological and economic implications for marine conservation and fisheries
management. As countries continue to implement LSMPAs as a way to
reach the stated 10\% target of ocean protection, managers should
consider how fishing effort will change in space and through time to
ensure that fishing effort is not just displaced elsewhere, leading to
overfishing in adjacent waters. While LSMPAs can provide a wide
range of benefits, their implementation must be accompained with
traditional fisheries management to maximize their effectiveness.

We do not observe an increase in daily fishing hours for vessels that used to fish within PIPA, compared to a control
group of vessels. Thus, for vessels that have chosen to continue to operate following the closure, we do not observe evidence that they are being forced to exert more effort as a result of the closure. Unfortunately, without detailed data on vessel catch, we cannot conclude whether vessels are better or worse off financially following the closure.

Our results suggest that PIPA has displaced some vessels to the high seas just outside the EEZ of Kiribati. This might imply a drop in demand for Kiribati fishing rights and a decline in revenue. We also uncover evidence that there is more congestion immediately following the closure, but this is declining over time.
\end{abstract}

@@

Next steps: I think we need to make the buffers around the EEZs 5 degrees instead of 1 degree (redistribution of effort section)
Might change the specification of the redistribution regression

\section{Introduction}\label{introduction}

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environment by
2020. In an effort to meet this target, there has been a rapid
increase recently in MPA coverage \citep{wood_2008,sala_2018}, largely driven by
a small number of Large Scale Marine Protected Areas (LSMPAs).\footnote{See, for example, \citet{game_2009},  \citet{singleton_2014},\citet{boonzaier_2016}, \citet{mccauley_2016}, and \citet{alger_2017}.}
). Today, a small number of LSMPAs make up at least 80\%
of the managed areas in the ocean \citep{toonen_2013}.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. As with smaller MPAs, it is important that we
understand the socioeconomic implications of management interventions.
One issue of particular importance is that of the displacement or
redistribution of fishing effort, which may influence the outcomes of an
MPA \citep{smith_2003}. Theoretical models make different assumptions
about the ways in which fishers will reallocate fishing effort after an
area closure, which often determine whether or not MPAs produce negative or positive impacts on fisheries. The existing empirical literature focuses on small MPAs and has been criticized for confounding correlation with causation \citep{ferraro2018causal}.

Recent advances in satellite tracking technologies and near real-time
identification of fishing activity provide us with an opportunity to ask, how do fishers respond to the implementation of an LSMPA? Does effort increase or decrease and where does it go? We use identification of fishing activity via Automatic
Identification Systems (AIS) and causal inference techniques to describe
the behavioral changes and spatial redistribution of the industrial tuna
purse seine fleet due to the implementation of the Phoenix Islands Protected Area. We use the same data to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary.

Our work is novel in the sense that it provides empirical evidence of
the effect of Large Scale Marine Protected Areas on fishing behavior and
distribution and can help guide future interventions. Understanding how
effort is displaced from LSMPAs might provide insights into how distant water fishing fleets would react to a high seas closure.

The paper is organized as follows: Section \ref{background} provides more
information on Large Scale Marine Protected Areas and some background on
empirical studies of effort redistribution. An overview of the Nauru
Agreement and associated countries, a description of the fleet that
operates in the region, and a brief history of PIPA and PNMS is also included.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, and Section \ref{discussion}
discusses our
results.

\section{Background}\label{background}

@@

The cutoff at which an MPA is considered to be a LSMPA ranges from areas
larger than 30,000 km\textsuperscript{2} as defined by
\citet{desanto_2013} or areas larger than 250,000 km\textsuperscript{2},
as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows
LSMPAs that meet the latter condition, and are also fully no-take. LSMPAs are often
implemented in the pelagic environment, where the dominant human
activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The
early literature on LSMPAs focused on the inherent challenges and
difficulties that come with a pelagic environment. \citet{kaplan_2010}
claimed that very large MPAs would result in excessive opportunity costs
and that these would be difficult to enforce. \citet{game_2009}
suggested that most of the challenges could be overcome with the
incorporation of technology, in what then became known as Dynamic Ocean
Management \citep{maxwell_2015}.\footnote{See \citet{singleton_2014}, who provide an objective discussion of the pros and cons of LSMPAs.}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}\label{fig:LSMPAs_map}Large Scale Marine
Protected Areas. The map shows all areas larger than 250,000
Km\textsuperscript{2}. Areas in dark blue meet at least one of these
conditions: reported no take area is greater than 0, are categorized as
IUCN Ia or Ib, their designated English name is `Protected Area'.}
\end{figure}

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citet{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protects two
thirds of the tagged grey reef sharks by effectively excluding fishing
effort. More recently, \citep{bradley_2018} use similar data to
highlight cases of potential illegal shark fishing \emph{inside} a 2
million km\textsuperscript{2} shark sanctuary. To date, no studies have
evaluated the displacement of fishing effort due to LSMPA
implementation.

Spatial closures of this magnitude are likely to induce changes in
fishers' behavior. Theoretical models of fishing effort redistribution
range from the simplistic assumption that effort inside the bounded
region disappears, to spatially explicit models that reallocate fishing
effort based on habitat characteristics, presence of other vessels, and
expected returns \citep{smith_2003,hilborn_2006}. However, these focus
on the long term optimal equilibrium, and redistribution of fishing
effort may not always be optimal, especially over the first few years
\citep{stevenson_2013}.

The empirical research that has been done in smaller sized MPAs
suggests that resource users may show idiosyncratic responses. For
example, \citet{stevenson_2013} show that a network of MPAs displaced
fishing effort farther away from ports, resulting in higher
\emph{perceived} costs, and increases in catch per unit effort.
\citet{cabral_2017} analyze the redistribution of fishing and
non-fishing vessels following the implementation of a network of MPAs in
California, and find that commercial dive boats (??????) follow a
fishing-the-line pattern, while some fishing boats follow an ideal free
distribution. More recently \citet{elahi_2018} used satellite tracking
data to show that a temporal spatial closure caused trawlers to maintain
effort but apply it more intensively elsewhere, particularly along the
borders and closer to shore. The way in which fishers react to a spatial
closure can have major implications for its outcome
\citep{smith_2003,hilborn_2006}, highlighting the need to understand how
fishers react to the implementation of LSMPAs, how fishing effort
changes, and how it is spatially redistributed. All these closures took
place within Exclusive Economic Zones, where other regulations exist.
This may not always be the case for LSMPAs, where often the entire EEZ
is converted into a LSMPA, leaving fishers with the option of moving to
the high seas or other countries' EEZs.

\hypertarget{nauru-agreement-and-the-phoenix-island-protected-area}{%
\subsection{Nauru agreement and the Phoenix Islands Protected
Area}\label{nauru-agreement-and-the-phoenix-island-protected-area}}

The Nauru Agreement was established in 1982 by a select group of Pacific
island nations to manage their important tuna resources more effectively. Parties to the Nauru Agreement (PNA) Members
include the Federated States of Micronesia, Kiribati, the Marshall Islands,
Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu. The Nauru
Agreement regulated access of foreign vessels (\emph{i.e.} those from
non-PNA countries). Holding \textasciitilde{}80\% of the historical
purse seining grounds within their Exclusive Economic Zones, PNA
countries gained greater bargaining power when providing fishing access to foreign
fleets \citep{havice_2010}.

The cooperation that emerged under the Nauru Agreement allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisers recommend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then use or sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\label{fig:PNA_map}Map of the Exclusive
Economic Zones (EEZs) of the region of interest. Countries that belong
to the PNA are shown in blue, while empty polygons indicates all others.
A red line indicates the Kiribati EEZ, and a solid red polygon
delineates PIPA.}
\end{figure}

@@

One of the most notable and recent management interventions in the
region is the implementation of the Phoenix Islands Protected Area (PIPA)
by the government of Kiribati. PIPA was first declared in 2006, and
established in 2008 with only 4\% of the area declared as no-take. On
January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
expanded to a total area of 397,447 km\textsuperscript{2}, roughly 1.5
times the size of Ecuador. Figure \ref{fig:PNA_map} shows a map of the
PNA countries and the Phoenix Islands Protected Area.

The closure of such a large area in one of the most important fishing
regions in the world provides a unique opportunity to evaluate the
behavioral responses and redistribution of fishing effort by vessels
that used to fish there. PIPA has been the focus of previous research
showing that fishing effort is effectively reduced after implementation
\citep{mccauley_2016,mcdermott_2018}. To this, we pose two questions:
How did individual vessels respond to the sudden exclusion of such a big
area? Where did all the vessels go? And what does this imply for revenues from
selling VDS for both Kiribati and the PNA as a whole?

A spatial closure might cause fishers to modify
their behavior as they adapt to a new state of the world. For example, some may have
to travel further distances to find new fishing grounds, increasing their fuel
costs. If fishers had developed experience for fishing in particular
sites, being excluded might impose a learning cost on them, as they
identify new fishing grounds. This might result in increased search
times. However, if the area of knowledge of a vessel is significantly
larger than that of the spatial closure, they might already know other
places to fish. In the next sections we describe
the data and methods used to answer these questions.

\section{Methods}\label{methods}

This section is divided into two main parts. First, we provide a general
description of AIS data and the process of identification of
vessel-level fishing events performed by Global Fishing Watch.\footnote{Global
	Fishing Watch: \url{globalfishingwatch.org}} Alongside, we describe
the subset of data used in our analysis. We also point
out possible shortcomings in the data, or factors that must be
considered in the analysis. We then move on to explain our
empirical strategy for the identification of behavioral changes and
the redistribution of fishing effort.

\subsection{Data}\label{data}

Automatic Identification Systems (AIS) are on-board devices that provide
at-sea safety and prevent ship collisions by broadcasting vessel
position, course, and activity to surrounding vessels. These broadcast
messages can be received by satellites and land-based antennas. GFW then uses 
machine learning techniques (convolutional neural networks) on the broadcast messages
to infer what type of fishing is taking place and where it is taking place, thus
allowing the estimation of near real-time fishing events globally \citep{kroodsma_2018}.

The amount of data gathered by GFW is dependent on the number of antennas and
satellites that can receive signals. The total satellite count increased from 3 to 6 on 
June 1\textsuperscript{st}
2014 , and then from 6 to 10 on January
1\textsuperscript{st} 2016. This causes an increase in the number of
\emph{received} AIS messages (\emph{i.e.} points), and therefore an
apparent increase in the number of vessels and fishing hours. However,
the addition of new satellites affects all vessels in the same way.

Our analysis focuses on purse seine vessels, the most important fishery
for PNA
countries.\footnote{We perform some of the same analyses for longliners and include them in the Appendix}
We identify a total of 103 purse seiners that fished in PNA waters at
least once before
2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses.}.
These vessels represent over 26 million individual observations for the
2012 - 2017 period. We identify 65 vessels that have fished inside PIPA
at least once since 2012. From these, 62 did so before the announcement
(\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}) but we are
only able to track 61 for the complete period of study before and after
its
implementation.\footnote{1 vessel never fished again after August 18, 2013.}
On the other hand, 38 vessels never fished inside PIPA but we only have
data for 26 of these vessels both before and after MPA implementation.
